---
title: "FML assignment 2"
output: html_document
date: "2023-10-01"
---

---
##Problem Statement

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of
relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In
particular, it wants to explore ways of converting its liability customers to personal loan
customers.
A campaign that the bank ran last year for liability customers showed a healthy conversion
rate of over 9% success. This has encouraged the retail marketing department to devise
smarter campaigns with better target marketing. The goal is to use k-NN to predict whether
a new customer will accept a loan offer. This will serve as the basis for the design of a new
campaign.
The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.
Partition the data into training (60%) and validation (40%) sets

```{r}
#Install required packages 
library(caret) #To split the dataset in training, validation, and testing.
library(class) #For classification of data
library(e1071) #For easy implementation of SVM
```
```{r}
#load the data
 UniversalBank <- read.csv("C:\\Users\\vasav\\OneDrive\\Desktop\\UniversalBank.csv")
 dim(UniversalBank)
 t(t(names(UniversalBank)))
```
Drop ID and ZIP 

```{R}
UniversalBank = UniversalBank[,-c(1,5)]
head(UniversalBank)
```

conversion of factor 

```{R}
#Only Education needs to be converted into Factor 
UniversalBank$Education = as.factor(UniversalBank$Education)
levels(UniversalBank$Education)

#Now, Convert Education to Dummy Variables

groups = dummyVars(~.,data = UniversalBank) #This created a dummy varible 

UniversalBank.Mod = as.data.frame(predict(groups,UniversalBank))

```

```{R}
set.seed(1) # Important to ensure that we get the same sample if we rerun the code

training.dif = sample(row.names(UniversalBank.Mod),0.6*dim(UniversalBank.Mod)[1])
validation.dif = setdiff(row.names(UniversalBank.Mod),training.dif)
train.diff = UniversalBank.Mod[training.dif,]
valid.diff = UniversalBank.Mod[validation.dif,]
t(t(names(train.diff)))

#Second approach

library(caTools)
set.seed(1)
split <- sample.split(UniversalBank.Mod, SplitRatio = 0.6)
train_set <- subset(UniversalBank.Mod, split == TRUE)
valid_set <- subset(UniversalBank.Mod, split == FALSE)

# Print the sizes of the training and validation sets
print(paste("The size of the training set is:", nrow(train_set)))
print(paste("The size of the validation set is:", nrow(valid_set)))
```

#Now, let us normalize data

```{R}
train.normal.diff <- train.diff[,-10] # Note that Personal Income is the 10th variable
valid.normal.diff <- valid.diff[,-10]

normal.values <- preProcess(train.diff[, -10], method=c("center", "scale"))
train.normal.diff <- predict(normal.values, train.diff[, -10])
valid.normal.diff <- predict(normal.values, valid.diff[, -10])

```

### Now let's solve the given Questions

#Question No:1

Consider the following customer:

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?

```{r}
# We have converted all categorical variables to dummy variables
# Let's create a new sample
New_Customer1 <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer
New.Cust.normal <- New_Customer1
New.Cust.normal <- predict(normal.values, New.Cust.normal)

```

#Now, let us predict using KNN

```{r}

KNN.Predct1 <- class::knn(train = train.normal.diff, 
                       test = New.Cust.normal, 
                       cl = train.diff$Personal.Loan, k = 1)
KNN.Predct1

```

***

2. What is a choice of K that balances between over-fitting and ignoring the predictor information?

```{r}
#Calculate the accuracy for each value of k
#Set the range of k values to consider

accuracy.diff <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  KNN.Predct <- class::knn(train = train.normal.diff, 
                         test = valid.normal.diff, 
                         cl = train.diff$Personal.Loan, k = i)
  accuracy.diff[i, 2] <- confusionMatrix(KNN.Predct,   
                                    as.factor(valid.diff$Personal.Loan),positive = "1")$overall[1]
}

which(accuracy.diff[,2] == max(accuracy.diff[,2])) 

plot(accuracy.diff$k,accuracy.diff$overallaccuracy)

```



***
3. Show the confusion matrix for the validation data that results from using the best k

```{R}

KNN.Predct2 <- class::knn(train = train.normal.diff, 
                         test = valid.normal.diff, 
                         cl = train.diff$Personal.Loan, k = 3)

confusionMatrix(KNN.Predct2,as.factor(valid.diff$Personal.Loan))

```
4.Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.

```{r}

#Classifying the customer using the best K.

New_Customer2 = data.frame(
  Age = 40, 
  Experience = 10, 
  Income = 84, 
  Family = 2,
  CCAvg = 2, 
  Education.1 = 0, 
  Education.2 = 1, 
  Education.3 = 0, 
  Mortgage = 0, 
  Securities.Account = 0, 
  CD.Account = 0, 
  Online = 1, 
  CreditCard = 1
)

KNN.Predct3 <- class::knn(train = train.normal.diff, 
                         test = New_Customer2, 
                         cl = train.diff$Personal.Loan, k = 3)

KNN.Predct3

#The customer has been classified as approved for personal loan

```

5.Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply
the k-NN method with the k chosen above. Compare the confusion matrix of the test set
with that of the training and validation sets. Comment on the differences and their reason.

```{R}
set.seed(2)
#Let's take 50% of the entire modified data as Training data 
train.diff2 = sample(row.names(UniversalBank.Mod), 0.5*dim(UniversalBank.Mod)[1])

#Let's take 30% of the data from the remaining 50% as Validation Data 
valid.diff2 = sample(setdiff(row.names(UniversalBank.Mod), train.diff2), 0.3*dim(UniversalBank.Mod)[1])

#Let's take remaining 20% of the modified data as Test Data
test.diff2 = setdiff(row.names(UniversalBank.Mod), union(train.diff2,valid.diff2))

train.normal.diff2 = UniversalBank.Mod[train.diff2,]
valid.normal.diff2 = UniversalBank.Mod[valid.diff2,]
test.normal.diff2 = UniversalBank.Mod[test.diff2,]

#transporting the data
t(t(names(train.normal.diff2)))


# Applying the k-NN method with the chosen K.

trainknn2 = knn(train = train.normal.diff2[,-8], test = train.normal.diff2[,-8], cl = train.normal.diff2[,8], k=3)

validknn2 = knn(train = train.normal.diff2[,-8], test = valid.normal.diff2[,-8], cl = train.normal.diff2[,8], k=3)

testknn2 = knn(train = train.normal.diff2[,-8], test = test.normal.diff2[,-8], cl = train.normal.diff2[,8], k=3)
```
Comparing the confusion matrix of the training set, validation sets and test set
```{r}

Confusionmatrix_trainknn2 = confusionMatrix(trainknn2, as.factor(train.normal.diff2$Personal.Loan),positive = "1")

Confusionmatrix_trainknn2


Confusionmatrix_validknn2 = confusionMatrix(validknn2, as.factor(valid.normal.diff2$Personal.Loan),positive = "1")

Confusionmatrix_trainknn2


Confusionmatrix_testknn2 = confusionMatrix(testknn2, as.factor(test.normal.diff2$Personal.Loan),positive = "1")

Confusionmatrix_trainknn2
```











